\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-27 forwards-and-backwards.tex
.

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $r$, and $F$,
with $\meanq = 0$ and $\covq = I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + F w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $F$, and $r$:
let $\lambda(x)$ be a smooth, nonnegative function on $\IR^d$,
\comment{TODO Should we worry that we used prime for differentiation before?}
and let $r'(x, m) = \lambda(x) r(x, m)$ and $F'(x, m) = \lambda(x) F(x, m)$
(and, let $\gamma' = \gamma$).
Since $\mu = r \gamma - F/\theta$,
this corresponds to multiplying both establishment probabilities and death rates by $\lambda$.
Then the population with parameters $\gamma'$, $r'$, and $F'$
has the same stationary profile(s) as the original population.

@@

\comment{TODO I think we need to make the notation for $r(y, \rho_r*\eta(y))$ 
consistent with what we have (now) used elsewhere, namely $r(y,\eta)$?}
Now suppose instead that
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
First note that
    \begin{align}
	    \nonumber
    & |B_f^\theta| =
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x, \smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
 \label{eqn:q_bound_split}
	    &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y, \smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
\nonumber
	    &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

>>>>

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $r$, and $F$,
with $\meanq = 0$ and $\covq = I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + F w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $F$, and $r$:
let $\lambda(x)$ be a smooth, nonnegative function on $\IR^d$,
and let $\widetilde{r}(x, m) = \lambda(x) r(x, m)$ and $\widetilde{F}(x, m) = \lambda(x) F(x, m)$
(and, let $\widetilde{\gamma} = \gamma$).
Since $\mu = r \gamma - F/\theta$,
this corresponds to multiplying both establishment probabilities and death rates by $\lambda$.
Then the population with parameters $\widetilde{\gamma}$, $\widetilde{r}$, and $\widetilde{F}$
has the same stationary profile(s) as the original population.

@@

Now suppose instead that
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
First note that
\begin{align}
	    \nonumber
    & |B_f^\theta| =
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
 \label{eqn:q_bound_split}
	    &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
\nonumber
	    &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy) 
        \right| .
\end{align}
(Note the extra term introduced here,
$r(x,\smooth{r}\eta(y))$, has the two arguments to $r$
``at different locations'', contrary to the usual pattern.)
